---
title: "helpful hints"
author: "George Perrett"
date: "2/1/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Discrete vs Continuous 
There are many ways to identify Discrete vs Continuous variables. In my own work, I commonly use `length(unique(x))` to get a vector of unique values in a vector. Lets say I define a Discrete variable as one with less than 5 unique values. I could use the following code: 


```{r}
# x is a vector in a data.frame
# cut is the number of unique values that classify a varible as continuous
get_type <- function(x){
  # if the variable is not numeric return discrete 
  if(is.numeric(x) == FALSE){type <- 'discrete'} 
  
  # if the variable is numeric but has less than 5 unique values return discrete
  else if(length(unique(x))<5){type <- 'discrete'}
  
  # everything else mark continuous
  else{type <- 'continuous'}
  
  
  return(type)
}

get_type(mtcars$cyl)
```

**Note we can still use normal R functions as helpers within Shiny reactive functions!!**
See week 2 class example on GitHub for applied example. 

# Plots 

base R histogram: 
```{r}
hist(mtcars$mpg)
```

bar chart in base R: 

```{r}
library(ggplot2)
#diamonds data is in ggplot2
plot(diamonds$cut)
```




ggplot2 histogram: 
```{r}
library(ggplot2)
ggplot(data = mtcars, aes_string("mpg")) + 
  geom_histogram(bins = 5) + # manually set number of bins defualt is 30 
  theme_bw() # makes the background white
```

ggplot2 bar chart: 
```{r}
ggplot(data = diamonds, aes_string("cut")) + 
  geom_histogram(stat = 'count') + # add stat = count for histogram of a string
  theme_bw() # makes the background white
```

ggplot2 goruped histogram: 
```{r}
ggplot(data = diamonds, aes_string('price', fill = 'cut')) + # fill is used to descirbe the variable to group and color by 
  geom_histogram() 
  theme_bw() # makes the background white
```

Another way to group histograms is by `facet` this will make a grid for each group

```{r}

# a friendly shiny way 
ggplot(data = diamonds, aes_string('price')) + 
  geom_histogram() + 
  facet_wrap(~get('cut')) + # facet wrap makes the pannel get() is what allows 'cut' to be read as cut
  theme_bw() # makes the background white

# a less friendly shiny way
ggplot(data = diamonds, aes_string('price')) + 
  geom_histogram() + 
  facet_wrap(~cut) + # note cut is not in '' which is how an input$ would come in 
  theme_bw() 

```

# renderUI

Sometimes we want to have a UI input that only shows up for a certain subset of selected variables. 

We can actually generate UI code in the server with `renderUI` and read it into the ui.R with `uiOutput`

We will cover this more in class next week! However, i've added an example to class2_examples if it is helpful



